Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-16 paper/paper.tex
correction

\subsection{An example}
\label{example}
Let's consider an example to understand the benefits of hierarchical architecture. Let there be a TE control application running on $d$ controllers, with each controller operating on $d$ switches. Network state consists of current utilizations of all links. Whenever a new flow arrives, application choses a path with minimum maximum utilization and assigns it to the flow. Each time a path of length $l$ is assigned to the flow, there could be a maximum of $l*d$ link utilization updates and $l$ forwarding rule messages to be sent between controllers. Now, consider an architecture where controllers are arranged in hierarchical manner. There are $d+1$ controllers. The first $d$ controllers (called type 0) maintain $d$ switches each and acts as a switch to the last controller (called type 1), which in turn maintains the other $d$ controllers. Type 0 controllers keep track of link utilizations only for the links that connect it's children switches. Type 1 controller keep track of links that runs from switches belonging to one controller domain to another. Same TE control application runs on each controller. When a new flow arrives, a switch sends $packet_{in}$ message to its type 0 controller. Type 0 controller checks if the path of the flow belongs entirely to its underlying network or not. If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller. However, if the destination is outside its domain, type 0 controller sends a $packet_{in}$ message to its parent controller. Parent controller uses TE control application to assign a paths of type 0 controllers. Forwarding rules are sent to type 0 controllers, which in turn sets up forwarding rules in their network. If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates. This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance. We show experimentally that the proposed architecture significantly reduces communication overhead while only modestly degrading application performance in the next section.

>>>>

\subsection{An example}
\label{example}
Let's consider an example to understand the benefits of hierarchical architecture. Let there be a TE control application running on $d$ controllers, with each controller operating on $d$ switches. Network state consists of current utilizations of all links. Whenever a new flow arrives, application choses a path with minimum maximum utilization and assigns it to the flow. Each time a path of length $l$ is assigned to the flow, there could be a maximum of $l*d$ link utilization updates and $l$ forwarding rule messages to be sent between controllers. Now, consider an architecture where controllers are arranged in hierarchical manner. There are $d+1$ controllers. The first $d$ controllers (called type 0) maintain $d$ switches each and acts as a switch to the last controller (called type 1), which in turn maintains the other $d$ controllers. Type 0 controllers keep track of link utilizations only for the links that connect it's children switches. Type 1 controller keep track of links that runs from switches belonging to one controller domain to another. Same TE control application runs on each controller. When a new flow arrives, a switch sends $packet\_in$ message to its type 0 controller. Type 0 controller checks if the path of the flow belongs entirely to its underlying network or not. If yes, it uses TE application to assign the path to the flow and no updates are sent to any controller. However, if the destination is outside its domain, type 0 controller sends a $packet\_in$ message to its parent controller. Parent controller uses TE control application to assign a paths of type 0 controllers. Forwarding rules are sent to type 0 controllers, which in turn sets up forwarding rules in their network. If the path length is $l$, there will be a maximum of $l+1$ forwarding rule messages between controllers and no link utilization updates. However, the path assigned might not be the one with minimum maximum utilization. This significant reduction in communication overhead comes at the cost of sub-optimality in the application performance. We show experimentally that the proposed architecture significantly reduces communication overhead while only modestly degrading application performance in the next section.
